---
title: "Mapping Assignment"
author: "Yu Du"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

# Set up:
```{r setup, include=FALSE,warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
               warning = FALSE,
               message = FALSE)

library(tidyverse)
library(drat)
library(hurricaneexposuredata)

library(ggplot2)
library(tmap)
library(tmaptools)
library(grid)
library(sf)
library(sp)
library(spData)
library(csodata)
library(dplyr)

addRepo("geanders")
data("hurr_tracks")
data("rain")
data("county.fips")
```

# For Floyd-1999:
```{r}
## For Floyd-1999
# Create a sub-dataset for the storm Floyd-1999 with hurricane data
data("county.fips")
#county.fips
Floyd_hurr <- hurr_tracks %>% 
  filter(storm_id == "Floyd-1999")
```

```{r}
# Create a sub-dataset for the storm Floyd-1999 with rain data
Floyd_rain <- rain %>% 
  filter(storm_id == "Floyd-1999")
```


```{r, message=FALSE}
#head(Floyd_hurr)
#head(Floyd_rain)
library(maps)
# Since the rain data contains fips only, needs to convert this information into
# longitude and latitude information 
#head(county.fips)
# The county.fips data contains information for fips, but fips are integers
Floyd_rain$fips <- as.integer(Floyd_rain$fips)
# head(Floyd_rain)
# Now, the fips in both tables are in a consistent format. 
```

```{r, message=FALSE}
Floyd_rain <- Floyd_rain %>% 
  group_by(fips) %>% 
  summarize(precip = sum(precip), precip_max = sum(precip_max))
```

```{r, message=FALSE}
# Add the county information into our rain data set
Floyd_rain <- left_join(Floyd_rain, county.fips, by = "fips")
# A new row polyname is added at the end, needs to seperate into two columns
library(magrittr)
Floyd_rain %<>% separate(polyname, c("county1", "county2"), sep = ",")

# The Floyd_rain now contains the county information as well
```

```{r}
# Check all counties showed in the rain data
# unique(Floyd_rain$county1)
counties_f <- c("alabama", "arkansas", "connecticut", "delaware", "district of columbia",
                "florida","georgia", "illinois", "indiana","iowa", "kansas", "kentucky", 
                "louisiana","maine", "maryland","massachusetts", "michigan", "mississippi",
                "missouri","new hampshire", "new jersey", "new york", "north carolina", 
                "ohio", "oklahoma", "pennsylvania", "rhode island", "south carolina",
                "tennessee", "texas","vermont","virginia", "west virginia", "wisconsin")  
# The map_data function provides longitude and latitude information for conuties
# head(map_data("county"))
counties_floyd <- map_data("county", counties_f)
state_floyd <- map_data("state", counties_f)
```

```{r}
# Rename the rain data set for consistency
Floyd_rain <-Floyd_rain %>% 
  rename(region = county1, subregion = county2)
# Add the longitude and latitude information into our rain data set
Floyd_rain <- left_join(counties_floyd, Floyd_rain,
                          by = c("region", "subregion"))
```


```{r}
# Create different range groups that will be showed on the map based on the 
# precip level
Floyd_rain <- Floyd_rain %>% 
  mutate(`Rainfall (mm)` = cut(Floyd_rain$precip, 
                     breaks = c(0, 25, 50, 75, 100, 125, 150, 175, 200, 225),
                     include.lowest = TRUE))
Floyd_rain <- na.omit(Floyd_rain)
#Floyd_rain
```

# For Allison-2001:
```{r}
## For Allison-2001:
# Similar steps for Allison-2001
Allison_hurr <- hurr_tracks %>% 
  filter(storm_id == "Allison-2001")
Allison_rain <- rain %>% 
  filter(storm_id == "Allison-2001")
```

```{r, message=FALSE}
Allison_rain$fips <- as.integer(Allison_rain$fips)
Allison_rain <- Allison_rain %>% 
  group_by(fips) %>% 
  summarize(precip = sum(precip), precip_max = sum(precip_max))
Allison_rain <- left_join(Allison_rain, county.fips, by = "fips")
Allison_rain %<>% separate(polyname, c("county1", "county2"), sep = ",")
```

```{r}
#unique(Allison_rain$county1)
counties_a <- c("alabama", "arkansas", "connecticut", "delaware", "district of columbia",
                "florida","georgia", "illinois", "indiana","iowa", "kansas", "kentucky", 
                "louisiana","maine", "maryland","massachusetts", "michigan", "mississippi",
                "missouri","new hampshire", "new jersey", "new york", "north carolina", 
                "ohio", "oklahoma", "pennsylvania", "rhode island", "south carolina",
                "tennessee", "texas","vermont","virginia", "west virginia", "wisconsin")  
counties_allison <- map_data("county", counties_a)
state_allison <- map_data("state", counties_a)
```

```{r}
Allison_rain <-Allison_rain %>% 
  rename(region = county1, subregion = county2)
Allison_rain <- left_join(counties_allison, Allison_rain,
                          by = c("region", "subregion"))
```

```{r}
# Create 2 different groups that will be showed on the map based on the precip level
Allison_rain <- Allison_rain %>% 
  mutate(`Rain > 175mm` = ifelse(Allison_rain$precip<175, "unexposed", "exposed"))

Allison_rain <- na.omit(Allison_rain)

```

# Floyd-1999 map with ggplot2:
```{r}
ggplot() + geom_polygon(data = Floyd_rain, aes(x = long, y = lat, group = group, 
                                                 fill = `Rainfall (mm)`), 
                        color = "grey", size = 0.2, alpha = 1.6) + 
  geom_polygon(data = state_floyd, aes(x = long, y = lat, group = group),
                 color="black", fill="white", size = 0.2, alpha = 0.3) +
  geom_path(aes(x = Floyd_hurr$longitude, y = Floyd_hurr$latitude), color = "orangered4") +
      scale_fill_brewer(palette = "Blues") +
  ggtitle("Floyd-1999") +
  # Center the title
  theme(plot.title = element_text(hjust = 0.5))
```

# Allison-2001 with ggplot2:
```{r}
  ggplot() + geom_polygon(data = Allison_rain, aes(x = long, y = lat, group = group, 
                                                 fill = `Rain > 175mm`), 
                          color = "grey", size = 0.2, alpha = 1.6) +
  geom_polygon(data = state_allison, aes(x = long, y = lat, group = group),
                 color="black", fill="white", size = 0.2, alpha = 0.1) +
  geom_path(aes(x = Allison_hurr$longitude, y = Allison_hurr$latitude), color = "orangered4") +
  scale_fill_manual(values = c("navyblue", "white"))+ 
  ggtitle("Allison-2001") +
  theme(plot.title = element_text(hjust = 0.5))
```

# Floyd-1999 with tmap:
```{r}
#Set up for tmap:
colnames(county.fips)[2]='ID'
county_sf1=st_as_sf(map('county',plot=F,fill=T))
county_sf1<-left_join(county_sf1,county.fips, by="ID")
Fl<-right_join(county_sf1,Floyd_rain, by="fips")

#Then plot tmap:
Breaks<-c(0, 25, 50, 75, 100, 125, 150, 175, 200, 225)
Fline<-cbind(Floyd_hurr$longitude,Floyd_hurr$latitude)%>%Line()%>%
  Lines(ID='Floyd-1999')%>%list()%>%SpatialLines()
tm_shape(Fl)+
  tm_polygons("precip",palette="Blues",title="Rainfall(mm)",breaks=Breaks)+
  tm_shape(Fline)+tm_lines(col='orangered4')+tm_layout(main.title='Floyd-1999',main.title.position = c('center', 'top'))
```

# Allison-2001 with tmap:
```{r}
#Set up for tmap:
county_sf2=st_as_sf(map('county',plot=F,fill=T))

county_sf2=left_join(county_sf2,county.fips,'ID')

#Then plot tmap:
Allison_line<-cbind(Allison_hurr$longitude,Allison_hurr$latitude)%>%Line()%>%
  Lines(ID='Allison-2001')%>%list()%>%SpatialLines()
Al<-right_join(county_sf2,Allison_rain, by="fips")
tm_shape(Al)+
  tm_polygons("Rain > 175mm",palette=c("navyblue","white"),title="Rain > 175 mm")+
  tm_shape(Allison_line)+tm_lines(col='orangered4')+tm_layout(main.title='Allison-2001',main.title.position = c('center', 'top'))
```

References:


Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686,
  https://doi.org/10.21105/joss.01686

Dirk Eddelbuettel with contributions by Carl Boettiger, Neal Fultz, Sebastian Gibb, Colin Gillespie,
  Jan Górecki, Matt Jones, Thomas Leeper, Steven Pav, Jan Schulz, Christoph Stepper, Felix G.M. Ernst
  and Patrick Schratz. (2020). drat: 'Drat' R Archive Template. R package version 0.1.8.
  https://CRAN.R-project.org/package=drat

Anderson B, Schumacher A, Crosson W, Al-Hamdan M, Yan M, Ferreri J, Chen Z, Quiring S, Guikema S (2020).
_hurricaneexposuredata: Data Characterizing Exposure to Hurricanes in United States Counties_. R package
version 0.1.0, <URL: https://github.com/geanders/hurricaneexposuredata>.

 H. Wickham. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York, 2016.

Tennekes M (2018). “tmap: Thematic Maps in R.” _Journal of Statistical Software_, *84*(6), 1-39. doi:
10.18637/jss.v084.i06 (URL: https://doi.org/10.18637/jss.v084.i06).

 Martijn Tennekes (2020). tmaptools: Thematic Map Tools. R package version 3.1.
  https://CRAN.R-project.org/package=tmaptools

 R Core Team (2020). R: A language and environment for statistical computing. R Foundation for
  Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.
  
 Pebesma, E., 2018. Simple Features for R: Standardized Support for Spatial Vector Data. The R Journal
  10 (1), 439-446, https://doi.org/10.32614/RJ-2018-009

Pebesma, E.J., R.S. Bivand, 2005. Classes and methods for spatial data in R. R News 5 (2),
  https://cran.r-project.org/doc/Rnews/.

Roger S. Bivand, Edzer Pebesma, Virgilio Gomez-Rubio, 2013. Applied spatial data analysis with R,
  Second edition. Springer, NY. https://asdar-book.org/
  
 Roger Bivand, Jakub Nowosad and Robin Lovelace (2020). spData: Datasets for Spatial Analysis. R
  package version 0.3.8. https://CRAN.R-project.org/package=spData  
  
 Eoin Horgan, Conor Crowley and Mervyn O'Luing (2020). csodata: Download Data from the CSO 'StatBank'
  API. R package version 1.0.2. https://CRAN.R-project.org/package=csodata
 
 Hadley Wickham, Romain François, Lionel Henry and Kirill Müller (2020). dplyr: A Grammar of Data
  Manipulation. R package version 1.0.2. https://CRAN.R-project.org/package=dplyr
  
  